A new form of transcorrelated Hamiltonian inspired by range-separated DFT

### 2021-04-26 cusp_j/paper/paper_draft.tex
added rsdft_cipsi

% f(x) = (3*erf(mu*x) - 1)/(2*x) + 1.5*mu/sqpi * exp(-(mu*x)**2.) - 0.25 * (1 - erf(mu*x))**2.
\subsection{Proposal of an ad-hoc symmetrization}
Let us assume that we only take a non symmetric form 
\begin{equation}
 \tilde{u}(\bd{r_1},\bd{r_2}) = u(r_{12};\mur{1}), 
\end{equation}
and we will see afterwards how to symmetrize the Hamiltonian.
\subsubsection{Computation of $\bigg( \nabla_1  u(r_{12};\mur{1})\bigg)^2$ }
therefore, according to Eq. \eqref{eq:grad_mur_1} one has 
\begin{equation}
 \begin{aligned}
 \deriv{}{x_1}{} u(r_{12};\mur{1}) =& \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}}\big( x_1 - x_2) \\
                                   +&\frac{e^{-(\mu(\bri{1}) r_{12})^2}}{2 \sqrt{\pi} \mu(\bri{1})^2} \deriv{\mur{1}}{x_1}{}.  
 \end{aligned}
\end{equation}
Therefore, the computation of $\big( \deriv{}{x_1}{} u(r_{12};\mur{1}) \big)^2$ yields 
\begin{equation}
 \begin{aligned}
 \bigg( \deriv{}{x_1}{} u(r_{12};\mur{1}) \bigg)^2 = & \frac{\bigg(1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)\bigg)^2}{4 \big(r_{12}\big)^2}\big( x_1 - x_2)^2 \\
 + & \frac{e^{-2(\mu(\bri{1}) r_{12})^2}}{4 \pi \mu(\bri{1})^4} \bigg(\deriv{\mur{1}}{x_1}{} \bigg)^2 \\ 
 + & \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{r_{12}}\big( x_1 - x_2) \frac{e^{-(\mu(\bri{1}) r_{12})^2}}{2 \sqrt{\pi} \mu(\bri{1})^2} \deriv{\mur{1}}{x_1}{}.
 \end{aligned}
\end{equation}
Therefore, the computation of $\bigg( \nabla_1  u(r_{12};\mur{1})\bigg)^2$ yields 
\begin{equation}
 \begin{aligned}
  \bigg( \nabla_1  u(r_{12};\mur{1})\bigg)^2  = &\frac{\bigg(1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)\bigg)^2}{4} 
   +   \frac{e^{-2(\mu(\bri{1}) r_{12})^2}}{4 \pi \mu(\bri{1})^4} \bigg( \nabla_1  \mur{1}\bigg)^2 \\ 
   + & \nabla_1  \mur{1} \, \cdot \, \big( \br{}_1 - \br{}_2\big) \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{r_{12}} \frac{e^{-(\mu(\bri{1}) r_{12})^2}}{2 \sqrt{\pi}\mu(\bri{1})^2} 
 \end{aligned}
\end{equation}

\subsubsection{Computation of $\nabla_1^2  u(r_{12};\mur{1})$ }
One has to perform first 
\begin{equation}
 \begin{aligned}
  \deriv{}{x_1}{2} u(r_{12};\mur{1}) =  & \deriv{}{x_1}{}  \bigg(\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}}\big( x_1 - x_2) \bigg) \\ 
      + & \deriv{}{x_1}{}  \bigg(\frac{e^{-(\mu(\bri{1}) r_{12})^2}}{2 \sqrt{\pi} \mu(\bri{1})^2} \deriv{\mur{1}}{x_1}{} \bigg).
 \end{aligned}
\end{equation}
Let us begin with the first term 
\begin{equation}
 \begin{aligned}
 \deriv{}{x_1}{}  \bigg(\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}}\big( x_1 - x_2) \bigg)  = 
&  \big( x_1 - x_2 \big) \deriv{}{x_1}{} \bigg(\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} \bigg) \\
+ & \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}}.
 \end{aligned}
\end{equation}
Then on has that 
\begin{equation}
 \begin{aligned}
 \deriv{}{x_1}{} \bigg(\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} \bigg) &= 
 \deriv{}{r_{12}}{} \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} \deriv{r_{12}}{x_1}{} \\  
 & + \deriv{}{\mur{1}}{}\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} \deriv{\mur{1}}{x_1}{}.
 \end{aligned}
\end{equation}
The first term reads 
\begin{equation}
 \begin{aligned}
& \deriv{}{r_{12}}{} \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} \deriv{r_{12}}{x_1}{} = \\
& -\frac{(x_1 - x_2)}{\big(r_{12}\big)^2} \bigg( \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} + \frac{\mu(\br{}_1)}{\sqrt{\pi}} e^{-\big(\mu(\br{}_1) r_{12}\big)^2}\bigg).  
 \end{aligned}
\end{equation}

>>>>

% f(x) = (3*erf(mu*x) - 1)/(2*x) + 1.5*mu/sqpi * exp(-(mu*x)**2.) - 0.25 * (1 - erf(mu*x))**2.
\subsection{Proposal of different symmetrization}
Let us assume that we only take the following form as a Jastrow factor 
\begin{equation}
 J(\br_1,\br_2) = \frac{e^{ u(r_{12};\mur{1})} + e^{u(r_{12};\mur{2})}}{2}. 
\end{equation}

\subsubsection{Computation of $\bigg( \nabla_1  u(r_{12};\mur{1})\bigg)^2$ }
therefore, according to Eq. \eqref{eq:grad_mur_1} one has 
\begin{equation}
 \begin{aligned}
 \deriv{}{x_1}{} u(r_{12};\mur{1}) =& \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}}\big( x_1 - x_2) \\
                                   +&\frac{e^{-(\mu(\bri{1}) r_{12})^2}}{2 \sqrt{\pi} \mu(\bri{1})^2} \deriv{\mur{1}}{x_1}{}.  
 \end{aligned}
\end{equation}
Therefore, the computation of $\big( \deriv{}{x_1}{} u(r_{12};\mur{1}) \big)^2$ yields 
\begin{equation}
 \begin{aligned}
 \bigg( \deriv{}{x_1}{} u(r_{12};\mur{1}) \bigg)^2 = & \frac{\bigg(1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)\bigg)^2}{4 \big(r_{12}\big)^2}\big( x_1 - x_2)^2 \\
 + & \frac{e^{-2(\mu(\bri{1}) r_{12})^2}}{4 \pi \mu(\bri{1})^4} \bigg(\deriv{\mur{1}}{x_1}{} \bigg)^2 \\ 
 + & \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{r_{12}}\big( x_1 - x_2) \frac{e^{-(\mu(\bri{1}) r_{12})^2}}{2 \sqrt{\pi} \mu(\bri{1})^2} \deriv{\mur{1}}{x_1}{}.
 \end{aligned}
\end{equation}
Therefore, the computation of $\bigg( \nabla_1  u(r_{12};\mur{1})\bigg)^2$ yields 
\begin{equation}
 \begin{aligned}
  \bigg( \nabla_1  u(r_{12};\mur{1})\bigg)^2  = &\frac{\bigg(1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)\bigg)^2}{4} 
   +   \frac{e^{-2(\mu(\bri{1}) r_{12})^2}}{4 \pi \mu(\bri{1})^4} \bigg( \nabla_1  \mur{1}\bigg)^2 \\ 
   + & \nabla_1  \mur{1} \, \cdot \, \big( \br{}_1 - \br{}_2\big) \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{r_{12}} \frac{e^{-(\mu(\bri{1}) r_{12})^2}}{2 \sqrt{\pi}\mu(\bri{1})^2} 
 \end{aligned}
\end{equation}

\subsubsection{Computation of $\nabla_1^2  u(r_{12};\mur{1})$ }
One has to perform first 
\begin{equation}
 \begin{aligned}
  \deriv{}{x_1}{2} u(r_{12};\mur{1}) =  & \deriv{}{x_1}{}  \bigg(\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}}\big( x_1 - x_2) \bigg) \\ 
      + & \deriv{}{x_1}{}  \bigg(\frac{e^{-(\mu(\bri{1}) r_{12})^2}}{2 \sqrt{\pi} \mu(\bri{1})^2} \deriv{\mur{1}}{x_1}{} \bigg).
 \end{aligned}
\end{equation}
Let us begin with the first term 
\begin{equation}
 \label{eq:_d_x_11}
 \begin{aligned}
 \deriv{}{x_1}{}  \bigg(\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}}\big( x_1 - x_2) \bigg)  = 
&  \big( x_1 - x_2 \big) \deriv{}{x_1}{} \bigg(\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} \bigg) \\
+ & \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}}.
 \end{aligned}
\end{equation}
Then on has that 
\begin{equation}
 \begin{aligned}
 \deriv{}{x_1}{} \bigg(\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} \bigg) &= 
 \deriv{}{r_{12}}{} \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} \deriv{r_{12}}{x_1}{} \\  
 & + \deriv{}{\mur{1}}{}\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} \deriv{\mur{1}}{x_1}{}.
 \end{aligned}
\end{equation}
The first term reads 
\begin{equation}
 \begin{aligned}
& \deriv{}{r_{12}}{} \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} \deriv{r_{12}}{x_1}{} = \\
& -\frac{(x_1 - x_2)}{\big(r_{12}\big)^2} \bigg( \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} + \frac{\mu(\br{}_1)}{\sqrt{\pi}} e^{-\big(\mu(\br{}_1) r_{12}\big)^2}\bigg),  
 \end{aligned}
\end{equation}
and the second term reads
\begin{equation}
 \begin{aligned}
 \deriv{}{\mur{1}}{}\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} \deriv{\mur{1}}{x_1}{}  = \\
 - \deriv{\mur{1}}{x_1}{} \frac{e^{-\big(\mur{1} r_{12}\big)^2}}{\sqrt{\pi}},
 \end{aligned}
\end{equation}
and eventually 
\begin{equation}
 \begin{aligned}
& \deriv{}{x_1}{} \bigg(\frac{1 - \text{erf}\big(\mur{1} r_{12} \big)}{2 r_{12}} \bigg) = 
- \deriv{\mur{1}}{x_1}{} \frac{e^{-\big(\mur(1) r_{12}\big)^2}}{\sqrt{\pi}}\\
& -\frac{(x_1 - x_2)}{\big(r_{12}\big)^2} \bigg( \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} + \frac{\mu(\br{}_1)}{\sqrt{\pi}} e^{-\big(\mur{1} r_{12}\big)^2}\bigg)
 \end{aligned} 
\end{equation}
and therefore Eq. \eqref{eq:_d_x_11} becomes 
\begin{equation}
 \begin{aligned}
& \deriv{}{x_1}{}  \bigg(\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}}\big( x_1 - x_2) \bigg)  = 
\frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} \\
&- \deriv{\mur{1}}{x_1}{} \frac{(x_1 - x_2)e^{-\big(\mur{1} r_{12}\big)^2}}{\sqrt{\pi}}\\
& -\frac{(x_1 - x_2)^2}{\big(r_{12}\big)^2} \bigg( \frac{1 - \text{erf}\big(\mu(\bri{1}) r_{12} \big)}{2 r_{12}} + \frac{\mu(\br{}_1)}{\sqrt{\pi}} e^{-\big(\mur{1} r_{12}\big)^2}\bigg)
 \end{aligned}
\end{equation}
